\chapter{Introduction}\label{ch:introduction}
Recommender systems (RS) are widely used to combat information overload for users when accessing new information by recommending specific items that the user is likely to find compatible \cite{YouTubeNeural,IndustryPerspective}.
RS are conventionally defined through three categories: \textit{content-based}, \textit{collaborative filtering} and \textit{hybrid}.
\textit{Content-based} RS provide recommendations based on content information such as genre, \textit{collaborative filtering} recommend items based on similar users, and \textit{hybrid} combines the two prior approaches \cite{ContextSurvey2020}.
\textit{Context-Aware Recommender Systems} (CARS) exist as a subset of RS, aiming to incorporate contextual information into the process of recommendation.
Contextual information is intuitively useful for making recommendations, in a movie recommendation situation this context could be company, time, or mood.
For example, a user may not want to watch the same movies in the company of their significant other, as they would with their parents.
Likewise, time could be an important factor, since movies such as \textit{A Christmas Story} may be more relevant during the winter months than during summer.\\\
Generally, for a movie RS, the objective is to estimate a rating function $R: User \times Item \rightarrow Rating$ for the $(user, \ item)$ pairs that are not yet rated \cite[p.~191-227]{RecommenderHandbook2015}.
CARS extends this function to $R: User \times Item \times Context \rightarrow Rating$, taking the known contextual information associated with the application into account.
Contextual information can be classified into three categories: \textit{fully observable}, \textit{partially observable} and \textit{unobservable}.
Fully observable contextual information is available explicitly, partially observable defines that only some of the information is available, and unobservable defines that no contextual information is explicitly available.
The structure of contextual information can remain stable over time, defined as static, or change in some way over time, defined as dynamic.
This paper will explore the state-of-the-art within CARS in order to examine open problems.
\todo{Rev til forskellige sections når de bliver lavet og beskriv struktur}

\section{Examples}
You can also have examples in your document such as in example~\ref{ex:simple_example}.
\begin{example}{An Example of an Example}
  \label{ex:simple_example}
  Here is an example with some math
  \begin{equation}
    0 = \exp(i\pi)+1\ .
  \end{equation}
  You can adjust the colour and the line width in the {\tt macros.tex} file.
\end{example}

\section{How Does Sections, Subsections, and Subsections Look?}
Well, like this
\subsection{This is a Subsection}
and this
\subsubsection{This is a Subsubsection}
and this.

\paragraph{A Paragraph}
You can also use paragraph titles which look like this.

\subparagraph{A Subparagraph} Moreover, you can also use subparagraph titles which look like this\todo{Is it possible to add a subsubparagraph?}. They have a small indentation as opposed to the paragraph titles.

\todo[inline,color=green]{I think that a summary of this exciting chapter should be added.}
